---
title: "Homework 4"
author: "Change-this-to-your-name"
date: "2/18/2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Choose a dataset to plot. You can use:

* built-in datasets in R, or 
* fetch one from a repository like [#tidytuesday](https://github.com/rfordatascience/tidytuesday#datasets), or
* use the coronavirus dataset (look at covid19_vaccine.Rmd for reading in)

Make sure you use `update_dataset()` if you choose either of the coronavirus datasets (some tidy formatting doesn't work if you don't).

You can list R built-in datasets in a new tab with the following code. This code also loads `mtcars` into your environment as an example.
```{r list-dataset}
data() # opens in new tab

#choose a dataset
data(mtcars) # adds the data to your environment
ls()
```

### QUESTION 1:
* What dataset (name) are you using? ______________
* Describe the dataset: _________________________________

```{r QUESTION-1-read-in-data}
# if you're using a built-in, just do data(datasetname)
# if you're using an external datasource, install or read it in
# if you're using coronavirus data, paste in the code from covid19_vaccine.Rmd
```

----

### Question 2: Barplot
```{r create-a-ggplot-barplot}
# use ggplot to create a bar plot
```

----

### Question 3: Add a column to the "fill" argument of aes()
```{r barplot-with-fill}
# add a "fill" argument to the aesthetic
```

----

### Question 4: ggplot extentions
Choose and describe a new plotting method from [this blog)(https://mode.com/blog/r-ggplot-extension-packages/).

*Points just for trying* the extension on a new plot on your dataset.
```{r try-out-extension}
# try out a new ggfunction
```