---
title: "Data Manipulation and Visualization with COVID-19"
author: ""
date: "2/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

This is an R package created and curated buy Ramin Krispin and Jerret Byrnes and 
has been uploaded to the [CRAN repository](https://cran.r-project.org/web/packages/coronavirus/readme/README.html). 
Currently, it contains two data sets which contain data on COVID19 cases, deaths, 
vaccinations, populations and countries among other data. We will use the two 
datasets within this coronavirus package to practice some data manipulation so
we can set ourselves up for data visualization.


### Open the toolbox
Let's begin by installing the package and a few tools with the ***install.packages() & library()*** functions
```{r}
# Use quotes "" for installing packages
#install.packages("coronavirus")
# install.packages("tidyverse")
library(coronavirus)
update_dataset()
library(tidyverse)
```

### Let's take a look at the data
```{r}
head(coronavirus)
tail(coronavirus)


head(covid19_vaccine)
tail(covid19_vaccine)
str(covid19_vaccine)
str(coronavirus)
```

```{r}
coronavirus %>%
  select(country) %>%
  unique() %>%
  count()

covid19_vaccine %>%  
  select(country_region) %>%
  unique() %>%
  count()

coronavirus %>%
  select(cases) %>%
  count()

covid19_vaccine%>%
  select(people_fully_vaccinated) %>%
  count()

# covid19_vaccine %>%
#   filter(country_region == "Afghanistan")
```

```{r}
# coronavirus %>%
#   select(c(country, population, cases, type)) 
# 
# type <- coronavirus %>%
#   select(type)
# 
# covid19_vaccine %>%
#   select(province_state) %>%
#   !is.na()
```

